---
title: "Country details in mapping"
output: 
  rmarkdown::html_vignette:
    keep_md: true 
    toc: true
vignette: >
  %\VignetteIndexEntry{Country details in mapping}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---




The CRAN version can be loaded as follows:
```{r, message = FALSE}
library('mapping')
```

or the development version from GitHub:

```{r, message = FALSE, eval=FALSE}
remotes::install_github('dataallaround/mapping')
```


## World


The worldwide geographical aggregates nomenclature  are the following:

| Function argument | Aggregates | Level | 
------------|------------| -------------|
| "continent" | Continents | level 0 |
| "regions" | World regions | level 1 |
| "subregion" | World subregions | level 2 |
| "country" | Countries | level 3 |

where **level 0** is the largest unit (continent), and **level 4** the smallest unit.

Countries can be subdivided in other aggregates, based on type of economics or development

| Function argument | Type |  
------------|------------| 
| "type" | Classification based on the type of political entity | 
| "type_economy" | Classification based on the type of development | 
| "type_income" | Classification based on the type of income | 

and based on the belonging to some international organisation

| Function argument | Type |  
------------|------------| 
| "nato" | If is a NATO country | 
| "ocde" | If is a ICDE country | 


```{r, echo=FALSE}
DiagrammeR::grViz("
  digraph graph2 {
  
  graph [layout = dot, rankdir = LR]
  
  # node definitions with substituted label text
  node [shape = oval]
  a [label = 'continent']
  b [label = 'regions']
  c [label = 'subregion']
  d [label = 'country']

  
  d -> c -> b -> a
  }
  
  ",height = 100)
```


`Level` may have names expressed in different way. The package provides:

|  Names | Code |  
------------|------------| 
| Country names | `country` | 
| iso2 code | `iso2` | 
| iso3 code | `iso3` | 
| iso3 numeric | `iso3_numeric` |
| iso3 United Nations  | `iso3_un` |
| iso2 World Bank  | `iso2_wb` |
| Formal name  | `name_formal` |
| World Bank name  | `name_wb` |

This can be specified in `matchWith` argument, depending on the **ids** format in the input data. 


```{r}
wr_continent <- loadCoordWR(unit = "continent", res = "hi")
mappingWR(wr_continent)
```

```{r}
wr_regions <- loadCoordWR(unit = "region", res = "hi")
mappingWR(wr_regions)
```


```{r}
wr_subregion <- loadCoordWR(unit = "subregion", res = "hi")
mappingWR(wr_subregion)
```

```{r}
wr_country <- loadCoordWR(unit = "country", res = "hi")
mappingWR(wr_country)
```


```{r}
wr_nato <- loadCoordWR(unit = "nato", res = "hi")
mappingWR(wr_nato)
```



```{r}
wr_ocde <- loadCoordWR(unit = "ocde", res = "hi")
mappingWR(wr_ocde)
```




## European Union


The European geographical aggregates nomenclature, called NUTS (Nomenclature des unités territoriales statistiques)) are the following:

| Function argument | Aggregates | Level | 
------------|------------| -------------|
| "nuts0" | European countries | level 0 |
| "nuts1" | European divisions for each country| level 1 |
| "nuts2" | European regions for each country| level 2 |
| "nuts3" | European municipalities for each country| level 3 |

where **level 0** is the largest unit (country), and **level 4** the smallest unit. In addition Urban Audit ("urau") are also provided.

```{r, echo=FALSE}
DiagrammeR::grViz("
  digraph graph2 {
  
  graph [layout = dot, rankdir = LR]
  
  # node definitions with substituted label text
  node [shape = oval]
  a [label = 'nuts0']
  b [label = 'nuts1']
  c [label = 'nuts2']
  d [label = 'nuts3']

  
  d -> c -> b -> a
  }
  
  ",height = 100)
```

The diagram show the hierarchy used in the `EU()` to build the object to map. Smaller aggregate/levels have all the information of the bigger aggregates. For example, if we have "nuts2", we have also all the information until the country level, "nuts0".



Level may have names expressed in different ways. The package provides: 

|  Names | Code |  
------------|------------| 
| Nuts names | `nuts` | 
| Nuts id code | `id` | 
| iso2 coutry code | `iso2` | 
| iso3 country code | `iso3` |

This can be specified in `matchWith` argument,depending on the **ids** in the input data.  


```{r}
eu_level0 <- loadCoordEU(unit = "nuts0")
mappingEU(eu_level0)
```

```{r}
eu_level0 <- loadCoordEU(unit = "nuts1")
mappingEU(eu_level0)
```


```{r}
eu_level0 <- loadCoordEU(unit = "nuts2")
mappingEU(eu_level0)
```

```{r}
eu_level0 <- loadCoordEU(unit = "nuts3")
mappingEU(eu_level0)
```



## Italy


The Italian geographical aggregates nomenclature are the following:

| Function argument | Aggregates | Level | 
------------|------------| -------------|
| "none" | country | level 0 |
| "ripartizione" | Italian divisions | level 1 |
| "regione" | Italian regions | level 2 |
| "provincia" | Italian municipalities | level 3 |
| "comune" | Italian districts | level 4 |

where **level 0** is the largest unit (country), and **level 4** the smallest unit.

```{r, echo=FALSE}
DiagrammeR::grViz("
  digraph graph2 {
  
  graph [layout = dot, rankdir = LR]
  
  # node definitions with substituted label text
  node [shape = oval]
  a [label = 'country']
  b [label = 'ripartizione']
  c [label = 'regione']
  d [label = 'provincia']
  e [label = 'comune']

  
  e -> d -> c -> b -> a
  }
  
  ",height = 100)
```

The diagram show the hierarchy used in the `IT()` to build the object to map. Smaller aggregate/levels have all the information of the bigger aggregates. For example, if we have "comune", we have also all the information until the first level, "ripartizione".

```{r}
it_level0 <- loadCoordIT(unit = "none")
mappingIT(it_level0)
```

```{r}
it_level1 <- loadCoordIT(unit = "ripartizione")
mappingIT(it_level1)
```


```{r}
it_level2 <- loadCoordIT(unit = "regione")
mappingIT(it_level2)
```

```{r}
it_level3 <- loadCoordIT(unit = "provincia")
mappingIT(it_level3)
```


```{r}
it_level4 <- loadCoordIT(unit = "comune")
mappingIT(it_level4)
```

Note that, given the government rule, the "provincia" and district may change over the `year`.


## USA

The USA geographical aggregates nomenclature are the following:

| Function argument | Aggregates | Level | 
------------|------------| -------------|
| "country" | USA country | level 0 |
| "region" | USA regions | level 1 |
| "division" | USA division | level 2 |
| "state" | USA states | level 3 |
| "county" | USA counties | level 4 |
| "district" | USA districts | level 5 |
| district_county| USA district in each county | level 6|

where **level 0** is the largest unit (country), and **level 6** the smallest unit. In addition Urban Area ("urban_area"") are also provided.

```{r, echo=FALSE}
DiagrammeR::grViz("
  digraph graph2 {
  
  graph [layout = dot, rankdir = LR]
  
  # node definitions with substituted label text
  node [shape = oval]
  a [label = 'country']
  b [label = 'region']
  c [label = 'region']
  d [label = 'state']
  e [label = 'county']
  f [label = 'district']
  g [label = 'district_county']

  
  g -> f -> e -> d -> c -> b -> a
  }
  
  ",height = 100)
```

The diagram show the hierarchy used in the `US()` to build the object to map. Smaller aggregate/levels have all the information of the bigger aggregates. For example, if we have "county", we have also all the information until the states level, "state".

Level may have names expressed in different way. The package provides:

|  Names | Code |  
------------|------------| 
| Names | `name` | 
| id | `id` | 
| Numeric code | `number` | 

This can be specified in `matchWith` argument, depending on the **ids** in the input data.



```{r}
us_level0 <- loadCoordUS(unit = "country")
mappingUS(us_level0)
```

```{r}
us_level1 <- loadCoordUS(unit = "region")
mappingUS(us_level1, type = "interactive")
```

```{r}
us_level2 <- loadCoordUS(unit = "division")
mappingUS(us_level2, type = "interactive")
```

```{r}
us_level3 <- loadCoordUS(unit = "state")
mappingUS(us_level3, type = "interactive")
```

```{r}
us_level4 <- loadCoordUS(unit = "county")
mappingUS(us_level4, type = "interactive")
```

```{r}
us_level5 <- loadCoordUS(unit = "district")
mappingUS(us_level5, type = "interactive")
```

```{r}
us_level6 <- loadCoordUS(unit = "district_county")
mappingUS(us_level6, type = "interactive")
```



