---
title: "mapping Italy"
output: 
  rmarkdown::html_vignette:
    keep_md: true 
    toc: true
vignette: >
  %\VignetteIndexEntry{mapping Italy}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE, 
                      fig.width = 7, fig.height = 4, fig.align = "center", 
                      message = FALSE, warning = FALSE)
```

The CRAN version can be loaded as follows:
```{r, message = FALSE}
library('mapping')
```

or the development version from GitHub:

```{r, message = FALSE, eval=FALSE}
remotes::install_github('serafinialessio/mapping')
```


The Italian geographical aggregates nomenclature are the following:

| Function argument | Aggregates | Level | 
------------|------------| -------------|
| "none" | country | level 0 |
| "ripartizione" | Italian divisions | level 1 |
| "regione" | Italian regions | level 2 |
| "provincia" | Italian municipalities | level 3 |
| "comune" | Italian districts | level 4 |

where **level 0** is the largest unit (country), and **level 4** the smallest unit.

```{r, echo=FALSE}
DiagrammeR::grViz("
  digraph graph2 {
  
  graph [layout = dot, rankdir = LR]
  
  # node definitions with substituted label text
  node [shape = oval]
  a [label = 'country']
  b [label = 'ripartizione']
  c [label = 'regione']
  d [label = 'provincia']
  e [label = 'comune']

  
  e -> d -> c -> b -> a
  }
  
  ",height = 100)
```

The diagram show the hierarchy used in the `IT()` to build the object to map. Smaller aggregate/levels have all the information of the bigger aggregates. For example, if we have "comune", we have also all the information until the first level, "ripartizione".

```{r}
it_level0 <- loadCoordIT(unit = "none")
mappingIT(it_level0)
```

```{r}
it_level1 <- loadCoordIT(unit = "ripartizione")
mappingIT(it_level1)
```


```{r}
it_level2 <- loadCoordIT(unit = "regione")
mappingIT(it_level2)
```

```{r}
it_level3 <- loadCoordIT(unit = "provincia")
mappingIT(it_level3)
```


```{r}
it_level4 <- loadCoordIT(unit = "comune")
mappingIT(it_level4)
```

Note that, given the government rule, the "provincia" and district may change over the `year`.
